---
title: "fireSense_SizeFit"
author: "Jean Marchal (jean.d.marchal@gmail.com)"
date: "August 2017"
output:
  html_document: default
  pdf_document: default
---

# Overview
Fit statistical models of the fire size distribution. A tapered Pareto distribution is assumed. This distribution has three parameters: $a$, $\beta$ and $\theta$; $a$ is the lower truncation point and is assumed to be known $a$ priori, $\beta$ controls the rate of frequency decrease as the fire size increases, and $\theta$ governs the location of the exponential taper. This module can be used to relate $\beta$ and $\theta$ with environmental controls of the fire size distribution.


# Download the module
```{r download module, eval = FALSE, echo = TRUE}
library(SpaDES)

moduleName <- "fireSense_SizeFit"

spadesModulesDirectory <- tempdir() # Location where the module will be downloaded

downloadModule(moduleName, path = spadesModulesDirectory)
```

# Usage
## Module parameters
Name|Default|Description
----|:--------------|---------------------------------------------------------------------
`formula`|`list(beta = NA,`<br>&emsp;&emsp;&emsp;`theta = NA)`|a named list with two elements, 'beta' and 'theta', describing the model to be fitted. 'beta' and 'theta' should be formulas (see `?formula`).
`a`||lower truncation point $a$ of the tapered Pareto. The random variable $x$ take values on the interval $a <= x < Inf$. Values outside of this range are ignored with a warning.
`link`|`list(beta = "log",`<br>&emsp;&emsp;&emsp;`theta = "identity")`|a named list with two elements, 'beta' and 'theta', specifying link functions for $\beta$ and $\theta$ parameters of the tapered Pareto. These can be character strings or objects of class link-glm. For more additional details see `?family`.
`data`|`"dataFireSense_SizeFit"`|a character vector indicating the names of objects in the `simList` environment in which to look for variables present in the model formula. `data` objects should be data.frames. If variables are not found in `data` objects, they are searched in the `simList` environment.
`start`|`NULL`|optional named list with two elements, 'beta' and 'theta', specifying starting values for the coefficients to be estimated. Those are passed to `nlminb` and can be a single vector, or a list of vectors. In the latter case, only the best solution, that is, the one which minimizes the most the objective function, is kept.
`lb`|`NULL`|optional named list with two elements, 'beta' and 'theta', specifying lower bounds for the coefficients to be estimated. These must be finite and will be recycled if necessary to match `length(coefficients)`.
`ub`|`NULL`|optional named list with two elements, 'beta' and 'theta', specifying upper bounds for the coefficients to be estimated. These must be finite and will be recycled if necessary to match `length(coefficients)`.
`itermax`|`500`|integer defining the maximum number of iterations allowed (DEoptim optimizer).
`nTrials`|`500`|if start is not supplied, `nTrials` defines the number of trials, or searches, to be performed by the nlminb optimizer in order to find the best solution.
`nCores`|`1`|non-negative integer. Defines the number of logical cores to be used for parallel computation. The default value is 1, which disables parallel computing.
`trace`|`0`|non-negative integer. If > 0, tracing information on the progress of the optimization are printed every `trace` iteration. If parallel computing is enable, nlminb trace logs are written to disk. Logs files are prefixed with 'fireSense_FrequencyFit_trace' followed by the subprocess pid. Default is 0, which turns off tracing.
`nlminb.control`|`list(iter.max = 5e3L,`<br>&emsp;&emsp;&emsp;`eval.max=5e3L`)|optional list of control parameters to be passed to the `nlminb` optimizer. See `?nlminb`.
`initialRunTime`|`start(simList)`|when to start this module? By default, the start time of the simulation.
`intervalRunModule`|`NA`|optional. Interval between two runs of this module, expressed in units of simulation time.
|||

## Usage example
```{r module usage example, eval = FALSE}
library(SpaDES)

set.seed(123)

times <- list(start = 1, end = 1, timeunit = "year")
modules <- list("fireSense_SizeFit")
paths <- list(
  modulePath = spadesModulesDirectory
)

# Example of model formula
model <- list(beta = formula(size_fires ~ MonthlyDroughtCode_June + PcHardwood + PcDisturbed + PcWaterBodies + PcOther),
              theta = formula(size_fires ~ MonthlyDroughtCode_July + PcHardwood + PcDisturbed + PcWaterBodies + PcOther))

parameters <- list(
  fireSense_SizeFit = list(
    formula = model,
    data = "dataFireSense_SizeFit",
    a = 1 # Lower truncation point, in the unit of measure used to measure fire sizes
  )
)

# Define from where and how data will be loaded in the simList environment
inputs <- data.frame(
  objectName = "dataFireSense_SizeFit",
  file = "/path/to/data/inputs",
  fun = "readRDS",
  package = "base",
  loadTime = 1
)

sim <- simInit(
  times = times,
  modules = modules,
  params = parameters,
  paths = paths,
  inputs = inputs
)

sim <- spades(sim)
sim$fireSense_SizeFitted
```

# Events
Events are scheduled as follows:

- Module initiation
- Model fitting

# Data dependencies

## Input data
- **dataFireSense_SizeFit**: one or more data.frames in which to look for variables present in the model formula.

## Output data
- **fireSense_SizeFitted**: an object of class `fireSense_SizeFit`, i.e. a list containing the following elements:

    - formula (model formula)
    - a (lower truncation point)
    - link (model link functions)
    - coef (fitted coefficients)
    - se (standard errors)
    - LL (log-likelihood)
    - AIC (Akaike Information Criterion)
    - convergence
    - convergence diagnostic

# Links to other modules
This module could be used in association with the fireSense_SizePredict module.

